### 2021-11-26 make_data.Rmd
Add manuscript, datamaking and fitted models

---
title: "Make data files"
output: 
  github_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

---

```{r setup, include=FALSE}
library(tibble)
library(dplyr)
library(tidyr)
library(stringi)
library(lubridate)
library(intervals)
library(igraph)
requireNamespace("here")

knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE
)
```

Processing the files from `data-src` to produce data frames and igraph/network objects for analysis.

The exact way in which the `gl200210.Rdata` was created is not 100% recovered thus we create the data files based on `wgl.Rdata`. Thus some of the code below is not executed until the full pipeline is re-checked.

# Trade data

## From source files

> Code in this section is not run.

Load original data files to separate environments (I'm not sure what object names are and where):

```{r load-data, eval = FALSE}
adata <- new.env()
load(here::here("data-src", "data", "NBA_AnalysisData.RData"), envir=adata)
ls(adata)

tdata <- new.env()
load(here::here("data-src", "data", "NBA_Transactions.RData"), envir=tdata)
ls(tdata)

mdata <- new.env()
load(here::here("data", "NBA_Meta.RData"), envir=mdata)
ls(mdata)

# Add rowid for back checking
main.df <- tdata$main.df %>%
  as_tibble() %>%
  tibble::rowid_to_column(".rowid")
```

Load standings data for team and division names:

```{r load-standings, eval = FALSE}
standings <- readRDS(here::here("data", "standings.rds")) %>%
  mutate(
    team_name = map_chr(strsplit(full_team_name, " "), tail, 1),
    Team = recode(team_name, "SuperSonics"="Sonics") # For joining with `standings`
  )
```

Create edgelist

```{r tradedf, eval=FALSE}
trade.df <- main.df %>%
  as_tibble() %>%
  select(.rowid, Date, Team, Notes) %>%
  mutate( 
    # Trade has "trade with" in `Notes` field...
    is_trade = grepl("trade with ", Notes),
    # ... but does not have the following phrases
    has_cancelled = grepl("cancelled ", Notes),
    has_rescinded = grepl("rescinded ", Notes),
    has_nullified = grepl("nullified ", Notes),
    has_voided = grepl("voided ", Notes)
  ) %>%
  filter(is_trade) %>%
  filter_at(vars(starts_with("has_")), all_vars(!.)) %>%
  select(-starts_with("has_"), -is_trade) %>%
  mutate( 
    # Remove stuff in parentheses from `Notes`
    notes = gsub("\\s*\\([^\\)]+\\)","", Notes),
    # Extract unique words starting with capital letter or digit (prospective team names)
    partner = map(stri_extract_all_regex(
      notes,
      paste0('\\b', paste0(unique(standings$Team), collapse="|"), '\\b'),
    ), unique),
    # Count extracted names
    num_trade_partners = map_int(partner, length)
  ) %>%
  transmute(
    team1 = Team,
    team2 = partner,
    date = Date,
    id = .rowid,
  ) %>%
  unnest("team2")
```

## Load seasonal weighted igraph objects

List of seasonal weighted igraph objects should be created from the items above. At this moment we just load it and create edge and vertex dataframes from it.

```{r load-wgl}
load(here::here("data", "wgl.Rdata"))
```

## Vertex database

A data frame for `season` x `name` with:

- `name` -- time-constant "franchise name"
- `label` -- time-varying team name
- `div` -- division name
- `season`, `season_from`, `season_to` -- season dates

```{r vdb}
vdb <- graphlist %>%
  setNames(nm = vapply(graphlist, graph_attr, character(1), name = "season")) %>%
  lapply(igraph::as_data_frame, what = "vertices") %>%
  bind_rows(.id = "season") %>%
  as_tibble() %>%
  separate(season, into = c("season_from", "season_to"), remove = FALSE) %>%
  mutate(
    across(starts_with("season_"), as.integer)
  )
vdb
```

## Edge database

A data frame for `season` x `from` x `to` where `from < to` with:

- `from`, `to` -- franchise names
- `season`, `season_from`, `season_to` -- season dates
- `weight` -- number of trades between `from` and `to` in `season`

```{r edb}
edb <- graphlist %>%
  setNames(nm = vapply(graphlist, graph_attr, character(1), name = "season")) %>%
  lapply(igraph::as_data_frame, what = "edges") %>%
  bind_rows(.id = "season") %>%
  as_tibble() %>%
  separate(season, into = c("season_from", "season_to"), remove = FALSE) %>%
  mutate(
    across(starts_with("season_"), as.integer),
    .from = pmin(from, to),
    .to = pmax(from, to)
  ) %>%
  select(
    from = .from, 
    to = .to, 
    season, 
    season_from,
    season_to,
    weight
  )
edb
```

## Checks

Team names in `vdb` and `edb` match:

```{r check}
u_vdb_teams <- with(vdb, unique(name))
u_edb_teams <- with(edb, unique(c(from, to)))
stopifnot(all(u_vdb_teams %in% u_edb_teams))
stopifnot(all(u_edb_teams %in% u_vdb_teams))
```

## Save

```{r save}
saveRDS(vdb, file = here::here("data", "df-vdb.rds"))
saveRDS(edb, file = here::here("data", "df-edb.rds"))
```

<!-- Include folded session info -->
<details>
<summary>Session info</summary>
```{r session-info}
sessioninfo::session_info()
```
</details>