### 2021-11-29 make_data.Rmd
Finish data processing script

---
title: "Data processing"
output: 
  github_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(stringi)
library(lubridate)
library(intervals)
library(igraph)
requireNamespace("here")

knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE
)
```

Producing data frames and igraph objects for analysis.

# Team, franchises and divisions

Seasonal team and franchise names along with division memberships based on data on standings from https://www.basketball-reference.com.

```{r teams-franchises-divisions}
teams <- readRDS(here::here("data", "standings.rds")) %>%
  distinct(
    season_from,
    season_to,
    team = purrr::map_chr(strsplit(full_team_name, " "), tail, 1) |>
      recode("Orleans" = "Hornets"),
    division = gsub("Division", "", division) |> trimws()
  )

teams <- teams |>
  bind_rows(
    filter(teams, season_from == 2018) |>
      mutate(
        season_from = 2019,
        season_to = 2020
      )
  ) |>
  mutate(
    franchise = case_when(
      # Map season-teams to franchises
      season_from <= 1978 & team == "Braves" ~ "Clippers",
      season_from >= 2004 & season_from <= 2013 & team == "Bobcats" ~ "Hornets",
      team == "Bullets" & season_from <= 1996 ~ "Wizards",
      team == "Hornets" & season_from >= 2002 & season_from <= 2012 ~ "Pelicans",
      team == "SuperSonics" & season_from <= 2007 ~ "Thunder",
      TRUE ~ team
    ),
    season = paste0(season_from, "-", season_to)
  )
```

## Checking selected franchise histories

Franchises in columns, time in rows, team names in cells.

```{r selected-histories, R.options=list(knitr.kable.NA="")}
teams |>
  filter(franchise %in% c("Hornets", "Bobcats", "Pelicans", "Wizards", "Thunder", "Clippers")) |>
  select(team, franchise, season_from) |>
  tidyr::pivot_wider(names_from = franchise, values_from = team) |>
  arrange(season_from) |>
  knitr::kable()
```

# Data on trades

Process data on trades.

- Assign trades to seasons using April 1 as a cut off date.
- Columns with team names are called `franchise1` and `franchise2` as the names pretend to "season-invariant" names of the franchises, not actual team names used in during seasons.
- Fix some franchise names mislabeled in the original data.

```{r trades}
load(here::here("data-src", "data", "NBA_AnalysisData.RData"), 
     envir=NBA_AnalysisData <- new.env())

trades <- NBA_AnalysisData$trade.df_long |>
  as_tibble() |>
  select(-season) |>
  mutate(
    season_from = case_when(
      month(Date) >= 4 ~ as.numeric(year(Date)),
      month(Date) <= 3 ~ as.numeric(year(Date)) - 1,
      TRUE ~ as.numeric(NA)
    ),
    season_to = season_from + 1,
    season = paste0(season_from, "-", season_to)
  ) |>
  # Fix franchise/team names in trades data
  mutate(
    P1 = case_when(
      P1 == "Pelicans" & season_from >= 1988 & season_from <= 2001 ~ "Hornets",
      TRUE ~ P1
    ),
    P2 = case_when(
      P2 == "Pelicans" & season_from >= 1988 & season_from <= 2001 ~ "Hornets",
      TRUE ~ P2
    )
  ) |>
  mutate(
    franchise1 = pmin(P1, P2),
    franchise2 = pmax(P1, P2)
  ) |>
  select(-P1, -P2)

# Make sure of canonical order
stopifnot(with(trades, all(franchise1 < franchise2)))
```

## Checks

Are there any franchise names in trade data that we can't match to standings on franchise name and season (expecting 0 rows in both)?

```{r}
trades %>%
  distinct(season_from, franchise1) %>%
  anti_join(teams, by = c("franchise1" = "franchise", "season_from")) %>%
  print(n = Inf)
```

```{r}
trades %>%
  distinct(season_from, franchise2) %>%
  anti_join(teams, by = c("franchise2" = "franchise", "season_from")) %>%
  print(n = Inf)
```

# Seasonal networks

Create a list of igraph objects with edge attribute `weight` counting the trades:

```{r graphlist}
graphlist <- lapply(
  sort(unique(trades$season_from)),
  function(s) {
    # Select trades
    e <- trades %>%
      filter(season_from == s) %>%
      select(franchise1, franchise2)
    # Nodes
    n <- teams %>%
      filter(season_from == s) %>%
      select(franchise, team, division)
    g <- graph_from_data_frame(e, vertices = n, directed=FALSE)
    E(g)$weight <- count_multiple(g)
    g <- simplify(g, edge.attr.comb = list(weight = "min"))
    g$season <- paste0(s, "-", s+1)
    g
  }
)

saveRDS(graphlist, file = here::here("data", "igraph-list.rds"))
```

# Edge and node data frames

Create node and edge data frames.

## Vertex database

A data frame for `season` x `name` with:

- `name` -- season-invariant "franchise name"
- `team` -- team name
- `division` -- division name
- `season`, `season_from`, `season_to` -- season dates

```{r vdb}
vdb <- graphlist %>%
  setNames(nm = vapply(graphlist, graph_attr, character(1), name = "season")) %>%
  lapply(igraph::as_data_frame, what = "vertices") %>%
  bind_rows(.id = "season") %>%
  as_tibble() %>%
  separate(season, into = c("season_from", "season_to"), remove = FALSE) %>%
  mutate(
    across(starts_with("season_"), as.integer)
  )

vdb

saveRDS(vdb, file=here::here("data", "nodes.rds"))
```

## Edge database

A data frame for `season` x `from` x `to` where `from < to` with:

- `from`, `to` -- trading franchises' names
- `season`, `season_from`, `season_to` -- season dates
- `weight` -- number of trades between `from` and `to` in `season`

```{r edb}
edb <- graphlist %>%
  setNames(nm = vapply(graphlist, graph_attr, character(1), name = "season")) %>%
  lapply(igraph::as_data_frame, what = "edges") %>%
  bind_rows(.id = "season") %>%
  as_tibble() %>%
  separate(season, into = c("season_from", "season_to"), remove = FALSE) %>%
  mutate(
    across(starts_with("season_"), as.integer),
    .from = pmin(from, to),
    .to = pmax(from, to)
  ) %>%
  select(
    from = .from, 
    to = .to, 
    season, 
    season_from,
    season_to,
    weight
  )

edb

saveRDS(edb, file = here::here("data", "edges.rds"))
```

## Checks

Team names in `vdb` and `edb` match:

```{r check}
u_vdb_teams <- with(vdb, unique(name))
u_edb_teams <- with(edb, unique(c(from, to)))
stopifnot(all(u_vdb_teams %in% u_edb_teams))
stopifnot(all(u_edb_teams %in% u_vdb_teams))
```

<!-- Include folded session info -->
<details>
<summary>Session info</summary>
```{r session-info}
sessioninfo::session_info()
```
</details>