### 2022-01-06 ergm-seasonal-results.Rmd
Seasonal model results

---
title: "Valued ERGMs"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, cache=FALSE}
requireNamespace("metafor")
library(dplyr)
library(ergm.count)
knitr::opts_chunk$set(
  echo = FALSE
  )
```

```{r load-data}
graphlist <- readRDS(here::here("data", "igraph-list.rds"))
results <- readRDS("ergm-seasonal.rds")

models <- list(
  A = net ~ sum + nodefactor("division"),
  B = net ~ sum + nodefactor("division") + nodematch("division"),
  C = net ~ sum + nodefactor("division") + nodematch("division", diff=TRUE),
  D = net ~ sum + nodematch("division")
) %>% 
  tibble::enframe("model_name", "model_formula")

d <- tibble::tibble(
  season = vapply(graphlist, igraph::get.graph.attribute, character(1), "season"),
  graph = graphlist, # igraphs
  net = lapply(graphlist, intergraph::asNetwork) # networks
) %>%
  tidyr::crossing(models) %>%
  mutate(
    model_formula = purrr::map2(net, model_formula, ~ {
      environment(.y)$net <- .x
      .y
    }
    )
  )

op <- options(warn = -1)

db <- d %>%
  transmute(
    season, 
    season_start = as.numeric(substr(season, 1, 4)),
    model_name,
    results = lapply(results, broom::tidy)
  ) %>%
  tidyr::unnest(results)

options(op)

saveRDS(db, "ergm-seasonal-db.rds")
```

The models are:

```{r}
models %>%
  knitr::kable()
```

# All models

The specifications fit to each season separately. Each cell is `coefficient (standard error) p-value`.

```{r}
db %>%
  transmute(
    season,
    model_name,
    term,
    cell = glue::glue("{round(estimate, 3)} ({round(std.error, 3)}) {format.pval(p.value, 3)}")
  ) %>%
  tidyr::pivot_wider(
    one_of("season", "term"),
    names_from = "model_name", 
    values_from = "cell",
    values_fill = "-"
  ) %>%
  knitr::kable()
```

# Uniform homophily coefficients

I.e. `nodematch` terms from models B and D:

```{r B-D-nodematch}
db %>%
  filter(
    model_name %in% c("B", "D"),
    term == "nodematch.sum.division"
  ) %>%
  knitr::kable()

db %>%
  filter(model_name == "B", term == "nodematch.sum.division") %>%
  pull(estimate) %>%
  hist(main = "Values of nodematch from models B")
```

## Funnel plots

### Model B

```{r B-nodematch-funnel-1976-2004}
with(
  db %>%
    mutate(
      season_start = as.numeric(substr(season, 1, 4))
    ) %>%
    filter( 
      model_name == "B", 
      term == "nodematch.sum.division",
      season_start < 2004
    ),
  metafor::funnel(estimate, vi = std.error^2, main = "1976 - 2004")
)  
```

```{r B-nodematch-funnel-2004-2019}
with(
  db %>%
    mutate(
      season_start = as.numeric(substr(season, 1, 4))
    ) %>%
    filter( 
      model_name == "B", 
      term == "nodematch.sum.division",
      season_start >= 2004,
      season_start < 2019
    ),
  metafor::funnel(estimate, vi = std.error^2, main = "2004 - 2019")
)  
```

```{r B-nodematch-funnel-allinone}
with(
  db %>%
    mutate(
      season_start = as.numeric(substr(season, 1, 4))
    ) %>%
    filter( 
      model_name == "B", 
      term == "nodematch.sum.division",
      season_start < 2019
    ),
  metafor::funnel(estimate, vi = std.error^2, 
                  col = c("black", "red")[(season_start > 2004) + 1]
  )
)
legend(
  "topright",
  pch = 19,
  col = c("black", "red"),
  legend = c("1976 - 2004", "2004 - 2019")
)
```

### Model D

```{r D-nodematch-funnel-1976-2004}
with(
  db %>%
    mutate(
      season_start = as.numeric(substr(season, 1, 4))
    ) %>%
    filter( 
      model_name == "D", 
      term == "nodematch.sum.division",
      season_start < 2004
    ),
  metafor::funnel(estimate, vi = std.error^2, main = "1976 - 2004")
)  
```

```{r D-nodematch-funnel-2004-2019}
with(
  db %>%
    mutate(
      season_start = as.numeric(substr(season, 1, 4))
    ) %>%
    filter( 
      model_name == "D", 
      term == "nodematch.sum.division",
      season_start >= 2004,
      season_start < 2019
    ),
  metafor::funnel(estimate, vi = std.error^2, main = "2004 - 2019")
)  
```

```{r D-nodematch-funnel-allinone}
with(
  db %>%
    mutate(
      season_start = as.numeric(substr(season, 1, 4))
    ) %>%
    filter( 
      model_name == "D", 
      term == "nodematch.sum.division",
      season_start < 2019
    ),
  metafor::funnel(estimate, vi = std.error^2, 
                  col = c("black", "red")[(season_start > 2004) + 1]
  )
)
legend(
  "topright",
  pch = 19,
  col = c("black", "red"),
  legend = c("1976 - 2004", "2004 - 2019")
)
```